A Fully Abstract Game Semantics for Countable Nondeterminism

### 2018-01-29 cones/paper/cones.tex
Did more things.

If $\ia\emptyset M$ is a configuration with empty store, we call $M$ a \emph{closed term}.  
Let $T$ be an Idealized Algol type, and let $M,N\from T$ be closed terms.
Then we write $M\lmam N$ if for all compatible contexts $C[-]$ of ground type we have
\begin{gather*}
  C[M]\converges V \Rightarrow C[N] \converges V\\
  C[M]\mustconverge \Rightarrow C[N] \mustconverge
\end{gather*}

@@

Tracking divergences explicitly in this way requires some care when we compose strategies.  
Specifically, we need to be able to add new divergences into strategies when they arise through `infinite chattering' or \emph{livelock}.  
For example, the denotation of the term
\[
  M = \Y_{\nat\to\nat}(\lambda f.\lambda n.n;(f n))\,,
  \]
where $n;P$ is a shorthand for $\IfO n\;P\;P$, is given by a total strategy, without divergences: namely the strategy $\mu$ with plays of the form
\[
  \begin{array}{cc}
    \mathbf{\bN} & \mathbf{\bN} \\
    & q \\
    q & \\
    n_1 & \\
    q & \\
    n_2 & \\
    \vdots &
  \end{array}\,.
  \]
However, when we compose this strategy with any total strategy for $\bN$ on the left, we expect the resulting strategy to contain divergences, since the term $M \n$ diverges for any $\n$.

@@

Given an arena $A$, we define an \emph{infinite justified string} in the obvious way.  
We define $\bar{P_A}$ to be $P_A$ together with the set of all those infinite justified sequences that have all finite prefixes in $P_A$.

Let $A$ be a game.  
A \emph{strategy} $\sigma$ for $A$ is a pair $(T_\sigma,D_\sigma)$, where:
\begin{itemize}
  \item $T_\sigma$ is a non-empty prefix-closed subset of $\bar{P_A}$ such that if $s\in T_\sigma$ is a $P$-position and $sa\in P_A$ then $sa\in T_\sigma$.
  \item $D_\sigma\subset \bar{P_A}$ is a postfix-closed set of plays in $\bar{P_A}$ that either end with an $O$-move or are infinite.  
    We require $D_\sigma$ to obey the following rules:
    \begin{description}
      \item[Divergences come from plays] If $d\in D_\sigma$ then there exists some $s\prefix d$ such that $s\in T_\sigma\cap D_\sigma$.
      \item[Diverge-or-reply] If $s\in T_\sigma$ is an $O$-position, then either $s\in D_\sigma$ or $sa\in T_\sigma$ for some legal play $sa$.
      \item[Infinite positions are divergent] If $s\in T_\sigma$ is infinite, then $s\in D_\sigma$.
    \end{description}
\end{itemize}

@@

We need to impose one more condition on strategies:
\begin{definition}
  Let $\sigma$ be a strategy for a game $A$.  
  We say that $\sigma$ is \emph{complete} if $T_\sigma=\bar{T_\sigma}$; i.e., $T_\sigma$ contains an infinite position $s$ if it contains every finite prefix of $s$.  
\end{definition}

@@

\begin{proposition}
  Let $\sigma\from I \to A$ be a strategy for a game $A$ in $\G_{ND}$.
  Then we may write $\sigma$ as $\top_{\bN};\Det(\sigma)$, where $\Det(\sigma)\from \oc\bN\to A$ is a deterministic strategy and $\top_{\bN}$ is the strategy for $\oc\bN$ given by
  \begin{itemize}
    \item $T_{\top_{\bN}}=P_{\oc\bN}$.
    \item $D_{\top_{\bN}}$ is the set of infinite positions in $T_{\top_\bN}$.
  \end{itemize}
\end{proposition}

>>>>

If $\ia\emptyset M$ is a configuration with empty store, we call $\ia \emptyset M$ a \emph{closed term}.  
In this case, we shall write $M$ rather than $\ia\emptyset M$.
Let $T$ be an Idealized Algol type, and let $M,N\from T$ be closed terms.
Then we write $M\lmam N$ if for all compatible contexts $C[-]$ of ground type we have
\begin{gather*}
  C[M]\converges V \Rightarrow C[N] \converges V\\
  C[M]\mustconverge \Rightarrow C[N] \mustconverge
\end{gather*}

@@

Tracking divergences explicitly in this way requires some care when we compose strategies.  
Specifically, we need to be able to add new divergences into strategies when they arise through `infinite chattering' or \emph{livelock}.  
For example, the denotation of the term
\[
  M = \Y_{\nat\to\nat}(\lambda f.\lambda n.n;(f n))\,,
  \]
where $n;P$ is a shorthand for $\IfO n\;P\;P$, is given by a total strategy, without divergences: namely the strategy $\mu$ with plays of the form
\[
  \begin{array}{cc}
    \mathbf{\bN} & \mathbf{\bN} \\
    & q \\
    q & \\
    n_1 & \\
    q & \\
    n_2 & \\
    \vdots &
  \end{array}\,.
  \]
However, when we compose this strategy with any total strategy for $\bN$ on the left, we expect the resulting strategy to contain divergences, since $M \n$ diverges for any $\n$.

@@

Given an arena $A$, we define an \emph{infinite justified string} in the obvious way.  
We define $\overline{P_A}$ to be $P_A$ together with the set of all those infinite justified sequences that have all finite prefixes in $P_A$.

Let $A$ be a game.  
A \emph{strategy} $\sigma$ for $A$ is a pair $(T_\sigma,D_\sigma)$, where:
\begin{itemize}
  \item $T_\sigma$ is a non-empty prefix-closed subset of $\overline{P_A}$ such that if $s\in T_\sigma$ is a $P$-position and $sa\in P_A$ then $sa\in T_\sigma$.
  \item $D_\sigma\subset \overline{P_A}$ is a postfix-closed set of plays in $\overline{P_A}$ that either end with an $O$-move or are infinite.  
    We require $D_\sigma$ to obey the following rules:
    \begin{description}
      \item[Divergences come from plays] If $d\in D_\sigma$ then there exists some $s\prefix d$ such that $s\in T_\sigma\cap D_\sigma$.
      \item[Diverge-or-reply] If $s\in T_\sigma$ is an $O$-position, then either $s\in D_\sigma$ or $sa\in T_\sigma$ for some legal play $sa$.
      \item[Infinite positions are divergent] If $s\in T_\sigma$ is infinite, then $s\in D_\sigma$.
    \end{description}
\end{itemize}

@@

We need to impose one more condition on strategies:
\begin{definition}
  Let $\sigma$ be a strategy for a game $A$.  
  We say that $\sigma$ is \emph{complete} if $T_\sigma=\overline{T_\sigma}$; i.e., $T_\sigma$ contains an infinite position $s$ if it contains every finite prefix of $s$.  
\end{definition}

@@

\begin{proposition}
  Let $\sigma\from I \to A$ be a strategy for a game $A$ in $\G_{ND}$.
  Then we may write $\sigma$ as $\top_{\bN};\Det(\sigma)$, where $\Det(\sigma)\from \oc\bN\to A$ is a deterministic strategy and $\top_{\bN}$ is the strategy for $\oc\bN$ that is given by
  \begin{itemize}
    \item $T_{\top_{\bN}}=P_{\oc\bN}$.
    \item $D_{\top_{\bN}}$ is the set of infinite positions in $T_{\top_\bN}$.
  \end{itemize}
\end{proposition}
